---
title: "One Line Plot - Points and Lines"
author: "Walter Xie"
date: "09/02/2017"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r library, warning=FALSE}
library(gg1L)
#getwd()
```

## Scatter plot

Create random coordinates for 2 clusters:

```{r random2Clusters, warning=FALSE}
df.clusters <- random2Clusters()
df.clusters$labels <- rownames(df.clusters)
df.clusters[1:3,]
```

Scatter plots may have overlaps between data points and axises. <code>unclip.ggplot</code> can turns off clipping, but it returns a *gtable* object, so that either <code>pdf.gtplot</code> or <code>plotgt</code> will be used to handle the plot.

```{r ScatterPlot, warning=FALSE}
gg.plot <- ggScatterPlot(df.clusters, x.id="x", y.id="y", colour.id="group", shape.id="group", 
                         title="Clusters", palette="Set1", verbose=F)
# turns off clipping
g.table <- unclip.ggplot(gg.plot) 
plotgt(g.table)
```

### Some tricks

Add <code>expand_limits</code> after ggplot object to expend x and y axes:

```{r ScatterPlot1, warning=FALSE, message=FALSE}
gg.plot + expand_limits(x = c(-1, 6), y=c(-1, 10))
```


<code>ellipsed.id="group"</code> creates ellipses by groups, <code>text.id="labels"</code> defines the column for labels, <code>xintercept=0 and yintercept=0</code> sets the coordinate axes, and <code>text.data=subset(df.clusters, x > 3 & y > 6)</code> selectively labels the points at x > 3 and y > 6:

```{r ScatterPlot2, warning=FALSE}
ggScatterPlot(df.clusters, x.id="x", y.id="y", colour.id="group", ellipsed.id="group", text.id="labels", 
              text.data=subset(df.clusters, x > 3 & y > 6), xintercept=0, yintercept=0,  
              title="Clusters", palette="Set1", verbose=F)
```

<code>text.or.point</code> controls appearance of labels and points:

* if 1, then display the texts only; 
* if 2, then only points; 
* if 3, the default, then both the texts and points. 

For example, show only labels:

```{r ScatterPlot3, warning=FALSE, message=FALSE}
ggScatterPlot(df.clusters, x.id="x", y.id="y", colour.id="group", text.id="labels", 
              text.or.point=1, title="Clusters", palette="Set1", verbose=F)
```

<code>text.repel</code> uses *geom_text_repel* from *ggrepel* package to avoid overlapped texts. 
It is better than <code>text.avoid.overlap</code>, where the latter sometime removes overlapped texts.

```{r ScatterPlot4, warning=FALSE, message=FALSE}
ggScatterPlot(df.clusters, x.id="x", y.id="y", colour.id="group", text.id="labels", 
              text.or.point=1, text.repel=T, title="Clusters (text.repel=T)", 
              palette="Set1", verbose=F)
```

<code>text.avoid.overlap</code> uses *check_overlap* in *geom_text* from *ggplot2* package:

```{r ScatterPlot5, warning=FALSE, message=FALSE}
ggScatterPlot(df.clusters, x.id="x", y.id="y", colour.id="group", text.id="labels", 
              text.or.point=1, text.avoid.overlap=T, title="Clusters (text.avoid.overlap=T)", 
              palette="Set1", verbose=F)
```

## Non-metric multidimensional scaling

Load the example data *reads.phyla*, where *x* is one of genetic marks, and the numbers are eDNA sequences (reads) identified as the phylum listed in *y*:

```{r reads.phyla, warning=FALSE}
data(reads.phyla)
reads.phyla[1:2,]
```

Because the functions to compute NMDS *metaMDS*, PCoA *cmdscale*, and PCA *prcomp* take different input data, please make sure you give a valid input for different NMDS plots.

### NMDS using *metaMDS* 

The following Nonmetric Multidimensional Scaling (NMDS) plot illustrates how the phyla are different in term of their reads abundances in 6 markers, and the data frame <code>reads.phyla[,7, drop=F]</code> provides the groups used for colouring texts. 

```{r NMDS1, warning=FALSE, message=FALSE}
nmds <- gtNMDSPlot(reads.phyla[,1:6], reads.phyla[,7, drop=F], colour.id="TaxaGroup", 
                   text.or.point=2, verbose=F)
plotgt(nmds)
```

### PCoA using *cmdscale* 

It is a classical MDS and also known as principal coordinates analysis PCoA (Gower, 1966).

```{r PCoA1, warning=FALSE, message=FALSE}
require(vegan)
dist.cm <- vegdist(reads.phyla[,1:6], method="jaccard", binary=TRUE)
cmds <- gtPCoAPlot(dist.cm, reads.phyla[,7, drop=F], colour.id="TaxaGroup", 
                   text.or.point=1, text.avoid.overlap=T, verbose=F)
plotgt(cmds)
```

### PCA using *prcomp* 

<code>prcomp</code> rotates the input to analyse on columns, so that we have the plot of 6 markers whose differences are in term of the abundances of phyla, 

```{r PCA1, warning=FALSE, message=FALSE}
pca <- gtPCAPlot(reads.phyla[,1:6], text.or.point=3, text.size=5, text.repel=T, verbose=F)
plotgt(pca)
```

## Line with points

The basic tutorial of line with points is avaible in another tutorial [**Plot traces**](https://cdn.rawgit.com/walterxie/gg1L/master/tutorials/OneLinePlotTracerR.html) subsection.

### Rarefaction curves for multi-sample

Coming soon.


