---
title: "One Command Line Plot - Tracer R"
author: "Walter Xie"
date: "10/02/2017"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

One of the feature of [**ComMA**](http://github.com/walterxie/ComMA) package is to simplify [ggplot2](http://docs.ggplot2.org/current/) codes into one-line functions, and also to provide some other useful utility functions.

```{r library, warning=FALSE}
library(ComMA)
#getwd()
```

## Load a MCMC log

Load a MCMC log from *../data-raw/star.beast.log*, the default to move the 1st data column into *row.names*, but using <code>row.names=NULL</code> to automatically generate row names and keep the 1st column in the data object:

```{r mcmc.log}
mcmc.log <- readMCMCLog("../data-raw/star.beast.log")
names(mcmc.log)
```

```{r mcmc.log2}
mcmc.log$state <- as.double(rownames(mcmc.log))
# remove burnin 10% 
burnin <- nrow(mcmc.log) * 0.1
mcmc.log <- mcmc.log[burnin:nrow(mcmc.log),]
# too many points, take the first 100 after burnin
mcmc.log100 <- mcmc.log[1:100, c("TreeHeight.Species", "state")]
mcmc.log100[1:3,]
```

## Plot traces

Plot the trace of the parameter *TreeHeight.Species*:

```{r LineWithPoints1, warning=FALSE}
lwp <- ggLineWithPoints(mcmc.log100, x.id="state", y.id="TreeHeight.Species", verbose=F)
# turns off clipping
gt.lwp <- unclip.ggplot(lwp) 
plot(gt.lwp)
```

### Adjust line and points

Show only line:

```{r LineWithPoints2, warning=FALSE}
ggLineWithPoints(mcmc.log100, x.id="state", y.id="TreeHeight.Species", line.or.point=1, verbose=F)
```                 
                 
Show only points:

```{r LineWithPoints3, warning=FALSE}
ggLineWithPoints(mcmc.log100, x.id="state", y.id="TreeHeight.Species", line.or.point=2, 
                 point.size=1, verbose=F)
```

## Kernel density estimate

```{r mcmc.log3, message=FALSE}
require(reshape2)
df.melt <- melt(mcmc.log, id="state")
# rename column "variable" to "parameters"
colnames(df.melt)[which(names(df.melt) == "variable")] <- "parameters"
# 3 TreeHeight
df.TreeHeight <- df.melt[grep("TreeHeight", df.melt[,"parameters"]),]
unique(df.TreeHeight$parameters)
df.TreeHeight[1:3,]
```

KDE for 3 parameters having the "TreeHeight" keyword:

```{r DensityEstimate1, warning=FALSE}
ggDensityEstimate(df.TreeHeight, x.id="value", colour.id="parameters", verbose=F)
```

Fill in colour:

```{r DensityEstimate2, warning=FALSE}
ggDensityEstimate(df.TreeHeight, x.id="value", fill.id="parameters", colour.id="parameters", verbose=F)
```

Stacked density plot to lose marginal densities:

```{r DensityEstimate3, warning=FALSE}
ggDensityEstimate(df.TreeHeight, x.id="value", fill.id="parameters", density.pos="stack", density.alpha=1, verbose=F)
```

Conditional density plot to preserve marginal densities:

```{r DensityEstimate4, warning=FALSE}
ggDensityEstimate(df.TreeHeight, x.id="value", y.id="..count..", fill.id="parameters", density.pos="stack", density.alpha=1, verbose=F)
```                  
                  
Percentage scale:

```{r DensityEstimate5, warning=FALSE}
ggDensityEstimate(df.TreeHeight, x.id="value", fill.id="parameters", density.pos="fill", density.alpha=1, verbose=F)        
```                  
                  
                