---
title: "One Line Plot - Bar Chart"
author: "Walter Xie"
date: "16/02/2017"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r library, warning=FALSE}
library(gg1L)
#getwd()
```

## Bar chart

Load the example data *model.test* from [ComMA](https://github.com/walterxie/ComMA) package (We only use the example data, *gg1L* doesn't depend it), select "Mac" only in *OS* column, and display the first 3 rows:

```{r model.test, warning=FALSE}
library(ComMA)
data(model.test)
model.test.mac <- model.test[model.test$OS=="Mac",]
model.test.mac[1:3,]
```

Plot the performance score of each test in four GTR models, where the texts of x-axis were rotated, and also use <code>verbose=F</code> to block messages:

```{r bar1, warning=FALSE}
ggBarChart(model.test.mac, x.id="test", y.id="performance", fill.id="model", x.text.angle=90, verbose=F)
```

### Adjust position of bars

Change the plot to stack bar chart using <code>bar.pos="stack"</code>, and title to "Stack Bar Chart":

```{r stackbar1, warning=FALSE}
ggBarChart(model.test.mac, x.id="test", y.id="performance", fill.id="model", x.text.angle=90, 
           bar.pos="stack", title="Stack Bar Chart", verbose=F)
```

Use both <code>bar.pos="fill"</code> and <code>y.trans="per"</code> to create a typical percentage bar chart:

```{r perbar1, warning=FALSE}
ggBarChart(model.test.mac, x.id="test", y.id="performance", fill.id="model", x.text.angle=90, 
           bar.pos="fill", y.trans="per", title="Percentage Bar Chart", verbose=F)
```

## Add something to bars

### Add error bars

 using <code>aggregate</code> modified from [the code] (http://www.r-bloggers.com/building-barplots-with-error-bars/) in R Bloggers:

```{r summary1, warning=FALSE}
myData <- aggregate(model.test$performance, by=list(model=model.test$model,OS=model.test$OS), 
                    FUN=function(x) c(mean=mean(x), sd=sd(x), n=length(x)))
myData <- do.call(data.frame, myData)
myData$se <- myData$x.sd / sqrt(myData$x.n)
colnames(myData) <- c("model", "OS", "performance", "sd", "n", "se")
myData[1:3,]
```

Or <code>summarySE</code> copied from [Cookbook for R](http://www.cookbook-r.com/Manipulating_data/Summarizing_data/):

```{r summary2, warning=FALSE, message=F}
myData <- summarySE(model.test, measurevar="performance", groupvars=c("model", "OS"))
myData[1:3,]
```

Calculate the lower and upper of error bars, where *performance* column is the mean.

```{r meanerr, warning=FALSE}
lower=myData$performance - myData$se
upper=myData$performance + myData$se
```

Add the error bars to the bar chart using <code>+</code>:

```{r errbar1, warning=FALSE}
bc <- ggBarChart(myData, x.id="model", y.id="performance", fill.id="OS", y.lab="performance", title="", verbose=F) 
bc + ggAddErrorBars(lower=lower, upper=upper)
```

### Add values

Add their values (in the column "n") to the bars:

```{r barnum1, warning=FALSE}
bc + ggAddNumbers(label.id="N", hjust=0.5, vjust=2)
```

Add the calculation result to the bars, such as the mean of y values:

```{r barnum2, warning=FALSE}
bc + ggAddNumbersFun(fun.y.lab=performance, y.adj=0.9)
```

### Add lines

Add the red line to show <code>y=1</code>:

```{r barline1, warning=FALSE}
bc + ggAddLine(linetype = 2, yintercept = 1, color="red")
```

<!-- bug
```{r barline2, warning=FALSE}
bc + ggAddLine(smooth.method = "lm")
```
--> 

## More Types

Load the example data *reads.phyla*, where *x* is one of genetic marks, and the numbers are eDNA sequences (reads) identified as the phylum listed in *y*:

```{r reads.phyla}
data(reads.phyla)
reads.phyla[1:2,]
```

### Percentage bar

This is a brief version of percentage bar, which uses <code>melt.id="TaxaGroup"</code> to specify the column of groups:

```{r perbar2, warning=FALSE}
pbc <- ggPercentageBarChart(reads.phyla, melt.id="TaxaGroup", verbose=F)
pbc$gg.plot
```

### Abundance bar by group

Use <code>melt.id="phylum"</code> to determine the value of bars (the numbers of reads of each phylum), but to group data and colour the bars by "TaxaGroup". Use *width* and *height* in <code>pdf</code> to resize the plot. It can also use the estimated width and height from the function.

```{r abundbar1, warning=FALSE}
# add melt.id
reads.phyla$phylum <- rownames(reads.phyla)
ab <- ggGroupAbundanceBar(reads.phyla, melt.id="phylum", colour.id="TaxaGroup", legend.row=1, verbose=F)
ab$gg.plot
#pdf.ggplot(ab$gg.plot, fig.path="abundance-bar.pdf", width=ab$pdf.width, height=ab$pdf.height)
```

<!--- mv to CMUtils.R because ggYAcrossXBar depends on ComMA::cmYAcrossX
### Y across X bar chart

This bar chart introduced in the supplymentary data of <https://doi.org/10.1186/s13742-015-0086-1> shows both of the total value of *y* (the number of reads) and the frequency of *y* (the number of phyla) appeared and only appeared in the given number of columns (datasets) listed in x-axis.  

```{r y-xbar1, warning=FALSE}
ggYAcrossXBar(reads.phyla[,1:6], terms=c("phyla", "datasets", "reads"), title="", 
              x.lab="Number of markers crossed", y.lab="Number of phyla/reads", verbose=F)
```
--->





